---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r echo=TRUE}
activity_data<-read.csv("activity.csv", header=TRUE, sep=",",na.strings="NA")
summary(activity_data)
```


## What is mean total number of steps taken per day?
### Total number of steps taken per day
```{r echo=TRUE}
total_steps<-aggregate(steps~date, data=activity_data, sum)
hist(total_steps$steps, main="Total Number of Steps per Day", xlab="Total Number of Steps per Day")
```

### Mean and median of the total number of steps taken per day
```{r echo=TRUE}
mean(total_steps$steps,na.rm=TRUE)
median(total_steps$steps,na.rm=TRUE)
total_mean<-mean(total_steps$steps,na.rm=TRUE)
total_median<-median(total_steps$steps,na.rm=TRUE)
```
The mean of the total number of steps taken per day is `r total_mean` and the median is `r total_median`.

## What is the average daily activity pattern?
### Time series plot
```{r echo=TRUE}
avg_steps<-aggregate(steps~interval, data=activity_data, mean)
plot(avg_steps$interval, avg_steps$steps, type="l", xlab="Interval", ylab="Average Steps", col="Blue")
```

### The 5-minute interval with maximum number of steps
```{r echo=TRUE}
max_avgsteps<-max(avg_steps$steps)
avg_steps$interval[avg_steps$steps==max_avgsteps]
```

## Imputing missing values
### Total number of missing values in the dataset
```{r echo=TRUE}
total_missing<-length(activity_data[is.na(activity_data)])
```
Total number of missing values in the dataset is `r total_missing`.

### Impute the missing values and make a histogram
The missing values are imputed by the mean of the same 5-minute interval.
```{r echo=TRUE}
new_data<-activity_data
for(i in 1:dim(new_data)[1]) {
  if(is.na(new_data$steps[i])) {
    new_data[i,1]<-avg_steps[avg_steps$interval==new_data[i,3],2]
  }
}
total_steps2<-aggregate(steps~date, data=new_data, sum)
hist(total_steps2$steps, main="Total Number of Steps per Day", xlab="Total Number of Steps per Day")
```

### Mean and median of the total number of steps taken per day based on the imputed data
```{r echo=TRUE}
mean(total_steps2$steps,na.rm=TRUE)
median(total_steps2$steps,na.rm=TRUE)
```

### The impacts of data imputation on the mean and median of the total number of steps taken per day
```{r echo=TRUE}
mean_change<-mean(total_steps2$steps,na.rm=TRUE)-mean(total_steps$steps,na.rm=TRUE)
median_change<-median(total_steps2$steps,na.rm=TRUE)-median(total_steps$steps,na.rm=TRUE)
```
The change in the mean of the total number of steps taken per day is `r mean_change`, and the change in the median of the total number of steps taken per day is `r median_change`. 

## Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE, warning=FALSE}
library(ggplot2)
weekday<-c('Monday', 'Tuesday', 'Wednesday','Thursday', 'Friday')
new_data$dayofweek<-factor((weekdays(as.Date(new_data$date)) %in% weekday), levels=c(TRUE,FALSE),labels=c('Weekday','Weekend'))
avg_steps2<-aggregate(steps~interval+dayofweek, data=new_data, mean)
g<-ggplot(avg_steps2, aes(x=interval, y=steps, group=dayofweek))
g+geom_line()+facet_wrap(~dayofweek,ncol=1)+
  labs(x="Interval", y="Number of Steps")+theme(panel.background=element_rect(fill='white', colour='black'),strip.background=element_rect(fill='orange', colour='black'),panel.grid.major = element_blank(), panel.grid.minor = element_blank())
```

